\documentclass{standalone}
\usepackage{diplomski}

\begin{document}

\chapter{Results}
\pagenumbering{arabic}
\setcounter{page}\thestranica

% --------------------------------------

The laboratory setup described in chapter \ref{ch:setup} enables simple change of measurement fibres, i.e. sensing elements, as well as use of different photodetectors. Here, different sets of single-mode and multi-mode fibres were tested. Different PIN photodiodes with electrical amplifiers and APD diodes were tested, according to table \ref{table:photodiodes}. Additionally, different laser configurations and data processing algorithms were tested in the development process of the measurement system.

\section{Single-mode fibres}

The first setup included single-mode fibres, connected in a manner displayed in Figure \ref{fig:smf_cascade}.
%\label{fig:smf_cascade}
On the near side of the measurement and excitation equipment, a 2.3 km-long roll of G.652-compliant single-mode fibre was connected. Its far end was spliced to a 4.2 km-long roll of highly non-linear single-mode fibre, i.e. a fibre with high $\left|\beta_3\right|$ value. Such a fibre will yield a higher non-linear response to the incident optical signal. In other words, lower power will be required to obtain the same SNR in the backscattered signal when using a highly non-linear fibre. A step in the scattered signal profile is visible, as a result of different $\left|\beta_3\right|$ coefficients. Fibres were terminated with an FC terminator, designed to provide minimum reflection. \\

The first measurements were conducted with a simple data processing algorithm, which used only a classical averaging filter. Filtered Stokes and anti-Stokes reflection profiles are presented in Figure \ref{fig:classic_average_compare}, with an averaging time of 15 seconds in Figure \ref{fig:classic_average_compare_15sec} and 7 minutes in Figure \ref{fig:classic_average_compare_7min}.
%\label{fig:classic_average_compare}
%\label{fig:classic_average_compare_15sec}
%\label{fig:classic_average_compare_7min}
%30.03.
The SNR improvement with time is evident, although further measurements did not yield any better results. The averaging filter has probably reached its best result in terms of the final SNR, while it still remains a very slow solution for filtering the signals. The next step in data processing algorithm development was the implementation of the adaptive LMS filter. A comparison of results between the two filters is given in Figure \ref{fig:classic_lms_compare}.
%\label{fig:classic_lms_compare}
%30.03.
The presented results are obtained with an averaging time of 7 minutes. While the results are very comparative, certain features of the adaptive LMS algorithm need to be considered. Firstly, in the anti-Stokes graph, the adaptive LMS filter has managed to extract finer features of the scattered profile. While the SNR for the anti-Stokes profile remains roughly the same, higher spectral components that may be the result of a physically measured quantity of interest, have passed through the adaptive LMS filter. Also, it is important to note that the same SNR was achieved with the adaptive LMS filter much sooner than with the classical averaging filter. The graph presented in \ref{fig:classic_lms_compare}, displaying the data from the adaptive LMS filter, had converged to its final form in less than one minute, dramatically decreasing the required measurement time. \\

While still observing the scattering profiles, instead of the calculated temperature, one must estimate the maximum incident optical power that is still below the stimulated Raman scattering threshold. In all the measurements, the first laser EDFA had a constant pump power
\begin{equation} \label{eq:edfa_first_pump}
I_1 = \SI{400}{\milli \ampere} \textrm{.}
\end{equation}
Lower first pump power setting ensured small ASE, although this has been further filtered. The second pump power was swept, and the peak optical power measured. The qualitative rating of the waveform, as well as the measured temperature inaccuracy is presented in Table \ref{table:smf_power_sweep}.
\begin{table}[h]
	\centering
	\caption{Sweeping optical power with an SM fibre}
	\label{table:smf_power_sweep}
	\begin{tabular}{C{2.5cm}|C{2.5cm}|C{2cm}|C{5cm}}
		\textbf{Pump current [mA]} & \textbf{Pulse peak power [mW]} & \textbf{$\bm{\varDelta T}$ [K]} & \textbf{Qualitative rating} \\
		\hline \hline
		600 & 730 & 200 & Profile visible \\
		700 & 2300 & 75 & -- \\
		800 & 3000 & 30 & -- \\
		900 & 3920 & 20 & Visible peak at the end of the profile \\
		1000 & 5540 & 15 & -- \\
		1050 & 6565 & 10 & Highly expressed peak at the end of the profile
	\end{tabular}
\end{table}
Pulse duration was set to 50 ns, and the measurement time was under 1 minute. A number of observations can be made here. Firstly, it can be seen that the pump power and the optical power have an irregular connection for this laser. In fact, the laser lacks power regulations, leading to the once set optical power being often unrepeatable in configuration. With some finer measurements, it was observed that a setting of
\begin{equation}
I_{2\textrm{max}} = \SI{1010}{\milli \ampere}
\end{equation}
yielding an optical power of about
\begin{equation}
P_\textrm{max} = \SI{5800}{\milli \watt}
\end{equation}
was a repeatable maximum power setting before the effects of SRS become significant. \\

Next, the laser impulse width was swept in order to determine the effect this will have to the measurement inaccuracy. It was shown that the width itself has little effect on the inaccuracy of the measured value itself. It is, however, important to ensure the laser width is long enough that the photodetectors, with their limited bandwidth, faithfully represent the impulse waveform and amplitude. The laser's pulse width can be configured in steps of 2.5 ns, and a value of
\begin{equation}
T_{0\textrm{min}} = \SI{12.5}{\nano \second}
\end{equation}
was determined as the minimal required pulse width. Shorter pulse width provides a finer spatial resolution, according to \ref{eq:otdr_resolution}. Although, due to the filtering nature of the decimation procedure performed on the calculated temperature profile, selecting a lower resolution will in fact yield a \textit{smoother} temperature profile. This is further discussed in the multi-mode scenario, where more comparative measurements were made. For many future measurements, a pulse width of
\begin{equation}
T_0 = \SI{25}{\nano \second}
\end{equation}
was selected as a representative configuration. \\

The final configuration parameter is the laser pulse repetition rate. Intuitively, we wish to have a rate that is large enough to provide a sufficient number of measurements in a given time frame, so that the data averaging can be effective. For most of the measurements, the repetition rate was set to
\begin{equation}
f_R = \SI{1}{\kilo \hertz} \textrm{.}
\end{equation}
The data processing software is not this fast. The runtime of a single processing iteration is estimated to be around 10--15 ms. Normally, setting a larger repetition rate would not pose a problem, as the software would only process the data that is last acquired from the DAQ card. However, architecture of the DAQ system needs to be taken into account. As the National Instruments PXIe-5160 has its own acquisition buffer memory, it is capable of performing the data acquisition when triggered, even if a large amount of data is still buffered and not transferred to the LabVIEW environment. A considerable backlog could occur, and the software would continue running the data acquisition loop long after the laser output was turned off. On the other hand, some hanging of the data acquisition software was noticed when setting the repetition rate too high. This could be a result of DAQ card's FIFO memory being overflown while the first data was being transferred to the LabVIEW software. While fine-tuning the DAQ part of the software, the repetition rate of
\begin{equation}
f_R = \SI{100}{\hertz}
\end{equation}
was found to be sufficient for this purpose. \\


An example of the final measurement, at maximum allowable optical power is given in Figure \ref{fig:smf_final}.
%\label{smf:final}
%13.04.
It is clear that only the highly non-linear fibre yields results with a reasonable accuracy, measured by this example to be about 
\begin{equation}
\varDelta T_\textrm{SMF} \approx \SI{20}{\kelvin} \textrm{.}
\end{equation}
Further data processing is expected to be able to bring the accuracy down to about 10 K. Measurements were made by PDA10CS-EC photodetectors, with the gain setting set to 40 dB. Therefore, we may expect some rapid temperature changes not to be detected by the system, as this setting provides a very narrow bandwidth. However, the problem with this setup is the accuracy of the sensing element itself, and future use of single-mode fibres was abolished.

\section{Multi-mode fibres}

As the effective cross-section area of a multi-mode fibre is larger than that of a single-mode fibre, the former is expected to demonstrate better sensing performance and temperature accuracy for this system. With a requirement that the power density does not cross that used in single-mode fibres at highest optical levels, multi-mode fibres will inherently be able to withstand higher peak optical powers before reaching SRS threshold. \\

As described in chapter \ref{ch:setup}, a lot of the optical components were changed to ensure coupling of all the incident optical power, and capturing all the backscattered power from the measurement fibre. For the first multi-mode setup, a cascade of two G.651-compliant fibres was used. The first fibre was 2.3 km, and the second 1.2 km in length. The two fibres were initially connected by SC-APC connectors, and terminated by another SC-APC connector with no matched termination at the end. \\

Different photodetectors were used in this setup. First measurements were performed with PDA10CS-EC with switchable gain. Firstly, the 40 dB setting was used to find the maximum peak optical power that this system could use. Laser's first EDFA pump current was set as in \ref{eq:edfa_first_pump}, while the second EDFA pump current was swept. The qualitative rating of the waveform, and the measured temperature inaccuracy is presented in Table \ref{table:mmf_power_sweep}.
\begin{table}[h]
	\centering
	\caption{Sweeping optical power with an MM fibre}
	\label{table:mmf_power_sweep}
	\begin{tabular}{C{2.5cm}|C{2.5cm}|C{2cm}|C{5cm}}
		\textbf{Pump current [mA]} & \textbf{Pulse peak power [mW]} & \textbf{$\bm{\varDelta T}$ [K]} & \textbf{Qualitative rating} \\
		\hline \hline
		1000 & 5600 & >10 & Profile clearly visible \\
		1100 & 9600 & <10 & -- \\
		1200 & 13500 & 5--10 & -- \\
		1250 & 15400 & <5--8 & High reflection off the end of the fibres
	\end{tabular}
\end{table}

%slika
%19.4.
%2-03
%ukljuci stokes as i temp


% --------------------------------------

\setcounter{stranica}{\thepage}
\addtocounter{stranica}{1}

\end{document}